Combining living and fossil taxa into phylogenies: the missing data issue

### 2014-06-11 Manuscript/TEM_draft.tex
Draft version 0.1 (Thomas)

We compared the "missing-data" tree to the "True" and the "Best" tree for each chain.
For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).

@@

\begin{figure} % Modify the figure
\centering
\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
\caption{Tree comparison protocol.
From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
We then compared this tree with or without fossil/living taxa to the "best/true" tree.
In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
\label{Fig_Compare}
\end{figure}

\subsubsection{Statistical difference between tree topologies}
To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g. effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
We used the pairwise tree comparisons as factors (i.e. the "Best" tree vs. one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}). %link broken

@@

\section{Results}
%Do the check for L00F00C00
\subsection{Building the trees}
When generating the matrices using seqgen and rTraitDisc algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
Using Bayesian inference, all the chains converged with a significant effective sample size (ASSD $<$ 0.01 and ESS $>>$ 200 for each tree).
Good enough phylogenetic signal.

@@

%10/06 - Missing 8 ML Chains

\subsubsection{Effect of $M_L$}
In Maximum Likelihood framework, the topological recovery quickly drops with the amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
However, this negative effect of missing data seems to stabilize beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
The Robinson-Fould metric shows a lower Tree similarity metric than the Triples metric meaning that individual taxa position are generally more conserved (i.e. low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).

Howerver, topological recovery was bad in Bayesian, the tree similarity score was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
However, it seems that there is a small difference depending on the metric used.
For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases.
In other words, if living taxa are removed, clades are less conserved (see Table ~\ref{Ba_RF-ML_results}).
However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).

@@

Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
And in the same way as described above, there seems to be a small difference depending on the metric used.
There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).

\subsubsection{Effect of $M_C$}
The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
This decrease is however slowlier when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).

@@

%To sum up:
The ability of recovering the "best" tree's topology in maximum likelihood method is function of the amount of data missing.
The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
However, in Bayesian, topology is badly (if not at all) recovered disregarding the amount of data missing.

\begin{table}
\caption{Tree similarity values per parameter in ML framework}
\centering
\begin{tabular}{rrlccccc}
    \hline
    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
              & missing data &        &      & lower  & upper & lower  & upper \\
    \hline
    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
              &              & Triples        & 1 & NA  & NA & NA  & NA \\
              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
              &              & Triples        & 1 & NA  & NA & NA  & NA \\
              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
              &              & Triples        & 1 & NA  & NA & NA  & NA \\
              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
    \hline
\label{NTS_ML_results}
\end{tabular}
\end{table}

\begin{table}
\caption{Tree similarity values per parameter in Bayesian framework}
\centering
\begin{tabular}{rrlccccc}
    \hline
    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
              & missing data &      & lower  & upper & lower  & upper \\
    \hline
    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
    \hline
\label{NTS_BA_results}
\end{tabular}
\end{table}

@@

\subsection{Building the trees}
Hard to simulate morpho characters but see Pagel, April Wright, etc...

\subsection{Comparing topologies}
Hard to compare Bayesian tree posterior distributions topologies but gives a good idea.

\subsection{ML vs Bayesian}
Missing data is not a problem as long as random. However, in TEM, n.fossils*n.molecular data is missing and not random.
Non optimal criterion approach (i.e. Bayesian) worse because of the number of considered "less"-likely trees (flat maximum likelihood landscapes).
However, ML still good performances and can be used to fix topology.

\subsection{Effect of missing data}
Parameter blabla seems to be influencing the topology the most.
But in practice, $M_L$ is the only one realistically solvable. Go to museums!

What about the "best" trees with fairly bad bootstrap support values in ML?
This is probably due to the low number of characters used in the simulation.
The number of morphological characters ($10^2$) is in the same order of magnitudes of empirical studies (cite) but decreases unrealistically with the $M_C$ parameter. %Extract the mean number of characters per matrices from the 100 random ones
The number of molecular characters ($10^3$) is at least two orders of magnitude lower than in empirical studies that can reach $10^6$ characters (cite).

Two different matrices from the "True" tree gives two different phylogenetic interpretations (i.e. seqGen and ape performs badly at generating a data matrix from a tree)?
That seems to not really be a problem since we are interested in recovering topology with missing data, not assessing if phylogenetic matrices generation software performs correctly.
Also, the "noise" that might be created by this potential "bad" phylogenetic matrices generation software is probably diminished by the number of replicates (i.e. 50 chains).

Different phylogenetic signal between the morphological and the molecular matrix?
That is not really a problem since the phenotypic evolution is not necessary correlated to the genotypic one.
Also it is important to note that both information are not exactly the same.
The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).

What about homoplasy in the morphological matrix?
Here, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly
If one states that morphological characters accumulate through time in the same way as the majority of the molecular characters (neutral theory, cite) then, homoplasic characters are expected to appear randomly through time.
Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
After a certain amount of morphological characters, adding new ones adds homoplasy \citep{wagner2000}.

Difference between Robinson-Fould and Triples?
Robison-Fould is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies.
When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa.
It will penalise only trees where taxa get removed furthest from their original clade.
Looking at our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e. where does the fossil branch) than the exact conservation of clades.

What's the crucial factor?
Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples vs. Robinson-Fould), the parameter that affects the more the decrease in topological recovery if the number of missing living taxa in the morphological part of the matrix ($M_L$).
This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
Therefore, we advocate that an effort should be made in collecting all the available morphological information for living taxa in order to properly run Total Evidence analysis. %yep, we need data from living taxa.

>>>>

We compared the "missing-data" tree to the "Best" tree for each chain.
For the Maximum Likelihood trees we performed pairwise comparisons between the "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).

@@

\begin{figure} % Modify the figure
\centering
\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
\caption{Tree comparison protocol.
From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
We then compared this tree with or without fossil/living taxa to the "best" tree.
In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
\label{Fig_Compare}
\end{figure}

\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g. effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
We used the pairwise tree comparisons as factors (i.e. the "Best" tree vs. one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}). %link broken

@@

\section{Results}

\subsection{Building the trees}
When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
This allows us to test our simulations in a theoretical as well as practical framework (i.e. phylogenies with low support values are rarely published but are still common in early stages of projects). %REF??? - That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  . %This is something that I will had with the whole submatrices story. Basically, when you have all the fossil/molecular data missing, it's harder to build the tree, both in ML and in Bayesian. When you run the analysis without missing data, then it works way better (i.e. our REAL control test: can we simulate matrices in the first place).
Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).

@@

%10/06 - Missing 8 ML Chains
\begin{figure} 
\centering
\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-results-BW.pdf}
\caption{Effect of missing data on Tree similarity.
$M_L$ is the amount of missing living taxa with morphological data,
$M_F$ is the amount of missing data in the fossil record,
$M_C$ is the amount of missing morphological characters.
The first row represent the Normalised Tree Similarity index using the Robinson-Fould metric;
the second row represent the Normalised Tree Similarity index using the Triples metric.}
\label{Fig_Results}
\end{figure}

\subsubsection{Effect of $M_L$}
In Maximum Likelihood framework, the topological recovery (reflected by the tree similarity) quickly drops with the increasing amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
However, this negative effect of missing data is stabilized beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
The Robinson-Fould metric shows a lower Tree similarity index than the Triples metric meaning that individual taxa position are generally more conserved (i.e. low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).

However, topological recovery was bad in Bayesian, regardless the amount of missing data.
The tree similarity index was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
Even if the topological recovery is really low, we detected an effect of missing data on the topology depending on the used metric.
For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the tree similarity index was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).

@@

Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
In the same way as described above, there seems to be a small difference depending on the metric used.
There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).

\subsubsection{Effect of $M_C$}
The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
This decrease is however slower when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).

@@

%To sum up:
The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.

\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
\caption{Tree similarity values per parameter in ML framework}
\centering
\begin{tabular}{rrlccccc}
    \hline
    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
              & missing data &        &      & lower  & upper & lower  & upper \\
    \hline
    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
              &              & Triples        & 1 & NA  & NA & NA  & NA \\
              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
              &              & Triples        & 1 & NA  & NA & NA  & NA \\
              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
              &              & Triples        & 1 & NA  & NA & NA  & NA \\
              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
    \hline
\label{NTS_ML_results}
\end{tabular}
\end{table}

\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
\caption{Tree similarity values per parameter in Bayesian framework}
\centering
\begin{tabular}{rrlccccc}
    \hline
    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
              & missing data &      & lower  & upper & lower  & upper \\
    \hline
    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
    \hline
\label{NTS_BA_results}
\end{tabular}
\end{table}

@@

\subsection{Building the trees}
Simulating evolutionary history matrices still remains a big drawback in theoretical phylogenetics. %Hahahaha.... feck.
The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
Even though molecular characters evolution (and therefore simulation) only depends of a small number of parameters (i.e. the base frequencies and the substitution matrix), simulating molecular matrix with a strong evolutionary signal is still complicated when generating unrealistically small matrices.
For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.

Also morphological characters studies involve many potential statistical pitfalls (e.g. independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy. %REF probably April as something about it
\begin{enumerate}
\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
%The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
\end{enumerate}
Our simulation parameters both decrease the phylogenetic signal (difficulty to simulate morphological characters and size of the matrix) as well as it increases it (reduction of homoplasy in the morphological part of the matrix).
Therefore, these drawbacks seems to have only a minor impact on the main results of this study: the incapacity of recovering any topology in Bayesian inference.

\subsection{Comparing topologies}
Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
Because our global framework is to study how to include efficiently fossils into phylogenies, we chose metrics reflecting the most interesting aspect of this global question: where do individual taxa (i.e. the fossils) branch in the tree.
The Robinson-Fould \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favoured.
Also by using the Normalised Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalised the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).

The Robinson-Fould metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
It will penalise only trees where taxa get removed furthest from their original clade.
Regarding our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e. where does the fossil branch) than the exact conservation of clades.

%Need a transition!
